Joshua 2.0: A Toolkit for Parsing-Based Machine Translation \\with Syntax, Semirings, Discriminative Training and Other Goodies

### 2010-04-23 papers/wmt10/joshua-progress.tex
Edited Chris Dyer's section
git-svn-id: https://joshua.svn.sourceforge.net/svnroot/joshua/branches/publications@1633 0ae5e6b2-d358-4f09-a895-f82f13dd62a4

Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly for low-resource languages. We have implemented an XML interface that allows external modules to propose alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}. Processing that is separate from the MT engine can suggest translations for some set of source side words and phrases. The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights. Additionally the XML format allows constraints on the righthand side of SCFG rules, which allows constraints like forcing a particular span to be a NP.  We modified the chart-based decoder to support these constraints.

@@

In \joshua, we use a {\em hypergraph} (or {\em packed forest}) to compactly 
represent the exponentially many derivation trees generated by the decoder for a 
particular foreign input sentence.
Given a hypergraph, we may perform many atomic inference operations,
such as finding one-best or $k$-best translations, or computing expectations over the hypergraph.
For each such operation, we could implement a dedicated dynamic programming algorithm.
However, a more general framework to specify these algorithms is semiring-weighted parsing \cite{semiringparsing}.
We have implemented  the inside algorithm, the outside algorithm, and the inside-outside speedup described by \newcite{li-eisner:2009:EMNLP},
plut the first-order expectation semiring \cite{eisner-expectation-semiring} and its
second-order version \cite{li-eisner:2009:EMNLP}.  All of these use our newly implemented semiring framework.

@@

\section{Lattice Parsing}

The bottom-up parsing algorithm used to generate the translation hypergraph has been generalized to support translation from not only an input sentence but from a \emph{word lattice} with runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.  This enables the exploration of a distribution over input sentences during decoding, and the best translation is selected from among any of them.  This functionality is useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters.

>>>>

Integrating output from specialized modules (like transliterators, morphological analyzers, and modality translators) into the MT pipeline can improve translation performance, particularly for low-resource languages. We have implemented an XML interface that allows external modules to propose alternate translation rules (constraints) for a particular word span to the decoder \cite{PBML-2010-Joshua-modules}. Processing that is separate from the MT engine can suggest translations for some set of source side words and phrases. The XML format allows for both hard constraints, which must be used, and soft constraints, which compete with standard extracted translation rules, as well as specifying associated feature weights. Additionally the XML format allows constraints on the lefthand side of SCFG rules, which allows constraints like forcing a particular span to be a NP.  We modified Joshua's chart-based decoder to support these constraints.

@@

In \joshua, we use a hypergraph (or packed forest) to compactly 
represent the exponentially many derivation trees generated by the decoder for an input sentence.
Given a hypergraph, we may perform many atomic inference operations,
such as finding one-best or $k$-best translations, or computing expectations over the hypergraph.
For each such operation, we could implement a dedicated dynamic programming algorithm.
However, a more general framework to specify these algorithms is semiring-weighted parsing \cite{semiringparsing}.
We have implemented  the inside algorithm, the outside algorithm, and the inside-outside speedup described by \newcite{li-eisner:2009:EMNLP},
plut the first-order expectation semiring \cite{eisner-expectation-semiring} and its
second-order version \cite{li-eisner:2009:EMNLP}.  All of these use our newly implemented semiring framework.

@@

\section{Word Lattice Input}

We generalized the bottom-up parsing algorithm that generates the translation hypergraph so that it supports translation from word lattices instead of just sentences.  Our implementation has runtime and memory overhead that is proportional to the size of the lattice, rather than the number of paths in the lattice \cite{dyer-muresan-resnik:2008:ACLMain}.  This allows distribution over input sentences to be explored during decoding, and allows the best translation to be selected from among all of them.  This is especially useful when \joshua is used to translate the output of statistical preprocessing components, such as speech recognizers or word segmenters, which can encode their alternative analyses as confusion networks or lattices.
